Caribou Demographic Projection - interesting & informative title TBD

### 2022-12-07 analysis/paper/paper.Rmd
little edits to paper intro

---
title: "Caribou Demographi Projection - interesting & informative title TBD"
author:
  - name: Josie Hughes
    email: josie.hughes@ec.gc.ca
    institute: [ECCC]
    correspondence: true
  - name: Sarah Endicott
    email: sarah.endicott@ec.gc.ca
    institute: [ECCC]
    correspondence: false
  - name: Cheryl A Johnson
    email: cheryl-ann.johnson@ec.gc.ca
    institute: [ECCC]
    correspondence: false
institute:
  - ECCC: Environment and Climate Change Canada 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      number_sections: no
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: CaribouDemographicModel.bib
csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
editor_options: 
  chunk_output_type: inline
---

@@

In Mississa (and other places) available local data is old and limited. Not clear how indicative older data is of state of population. Need new data to resolve questions, but what type, how much is enough. Can't monitor in great detail everywhere all the time. More generally, need for tools to help assess the value of additional monitoring, given the info that is already available.

@@

* How get a demographic forecast informed by local demographic data and national disturbance-demographic relationships?
* How to assess the value of collecting additional demographic data? 
* How does that value depend on current status of the range, available existing demographic data, and projected future disturbance. When and why is more information required, and when do we know enough?
* Provide decision support tools to enable others to get forecasts informed by their local data and national models, and assess the value of monitoring scenarios.

@@

>>>>

---
title: "Caribou Demographic Projection - interesting & informative title TBD"
author:
  - name: Josie Hughes
    email: josie.hughes@ec.gc.ca
    institute: [ECCC]
    correspondence: true
  - name: Sarah Endicott
    email: sarah.endicott@ec.gc.ca
    institute: [ECCC]
    correspondence: false
  - name: Cheryl A Johnson
    email: cheryl-ann.johnson@ec.gc.ca
    institute: [ECCC]
    correspondence: false
institute:
  - ECCC: Environment and Climate Change Canada 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      number_sections: no
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: CaribouDemographicModel.bib
csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
editor_options: 
  chunk_output_type: inline
---

@@

Lots of variability among caribou ranges in monitoring data: recency, duration, types of data. E.g. 27+ consecutive years of both telemetry-survival and calf/cow-ratios in most Alberta ranges, vs. 1-4 years of data in most Ontario ranges (and that's old) *note we may want to be careful about inter-province comparisons for political reasons, but the variation is super relevant when to-date we've had the same 'rules' apply across the board...

In Missisa (and other places) available local data is old and limited. Not clear how indicative older data is of state of population. Need new data to resolve questions, but what type, how much is enough. Can't monitor in great detail everywhere all the time. More generally, need for tools to help assess the value of additional monitoring, given the info that is already available.

@@

* How get a demographic forecast informed by local demographic data and national disturbance-demographic relationships?
* How to assess the value of collecting additional demographic data? 
* maybe also something like "How much of a bias/misinterpretation can be (has been?) incurred if we don't recognize local differences"?
* How does that value depend on current status of the range, available existing demographic data, and projected future disturbance. When and why is more information required, and when do we know enough?
* Provide decision support tools to enable others to get forecasts informed by their local data and national models, and assess the value of monitoring scenarios.

@@

## Caribou population monitoring: survival and recruitment metrics

Add in some brief description of the traditional approaches to estimating caribou survival (telemetry) and recruitment (aerial surveys/calf-cow ratios) - and show how these have been key for estimating population growth and status in most major boreal caribou projections/status documents. This could be useful background for defining terms below e.g. Rt, ct, at
